Dark Control:\\ Towards a Unified Account of Default Mode Function\\ by Markov Decision Processes

### 2017-12-07 paper/paper.tex
minor edits

% official NIPS keywords
\textbf{\\keywords}: mind-wandering, artificial intelligence, Markov decision processes

@@

\section{Conclusion}
Which brain function could be important enough
for the existence and survival of the human species
to justify constantly high energy costs?
MDPs motivate an attractive
formal account of how the human association cortex
can be thought to implement multi-sensory representation and decision making to
optimize the organism's intervention on the world.
% by obtaining reward and avoiding pain.
This idealized process model accommodates
a number of previous observations from neuroscience experiments on the
DMN by simple but non-trivial mechanisms.
%
From the view of a Markovian sequential decision process,
human behavior unfolds by inferring cumulative reward outcomes from hypothetical action cascades
% stored past events
and extrapolation from past experience to
upcoming events for guiding behavior in the present.
MDPs also provide a mathematical formalism how
opportunity in the environment can be recursively exploited
when confronted with challenging decisions.
%
This functional interpretation may well be compatible with the DMN's
poorly understood involvement across
autobiographical memory recall, problem solving,
abstract reasoning, social cognition,
as well as delay discounting and self prospection into the future.
%
Improvement of the internal world representation
by injecting stochasticity into the recall of past
actions and the estimation of action outcomes may explain why
highly accurate memories have been disfavored in human evolution
and why human creativity may be adaptive.

A major hurdle in explaining DMN activity from brain-imaging studies has been its very similar engagement across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., daytime mind-wandering), and thinking about future scenarios (e.g., delay discounting). The MDP account of DMN function offers a natural integration of these a-priori diverging mental processes into a common framework.
%
It is an important feature of the proposed artificial intelligence perspective
on DMN biology that it is practically computable and
proposes neuroscientific hypotheses that can be put to the test in future research.
%
Neuroscience experiments on the DMN should be designed that operationalize
the set of action, value, and state variables governing
the behavior of intelligent RL agents.
%
At the least, we propose an alternative vocabulary to
describe, contextualize, and interpret experimental findings in neuroscience studies
on higher-level cognition.
%
Ultimately,
DMN activity may instantiate a holistic integration
ranging from real experience over purposeful dreams to anticipated futures
for continued refinement of the organism's fate.

>>>>

% official NIPS keywords
\textbf{\\keywords}: systems neuroscience, artificial intelligence, reinforcement learning, mind-wandering

@@

\section{Conclusion}
Which brain function could be important enough
for the existence and survival of the human species
to justify constantly high energy costs?
MDPs motivate an attractive
formal account of how the human association cortex
can be thought to implement multi-sensory representation and decision making to
optimize the organism's intervention on the world.
% by obtaining reward and avoiding pain.
This idealized process model accommodates
a number of previous observations from neuroscience studies on the
DMN by simple but non-trivial mechanisms.
%
From the view of a Markovian sequential decision process,
human behavior unfolds by inferring cumulative reward outcomes from hypothetical action cascades
% stored past events
and extrapolation from past experience to
upcoming events for guiding behavior in the present.
MDPs also provide a mathematical formalism how
opportunity in the environment can be recursively exploited
when confronted with challenging interdependent decisions.
%
This functional interpretation may well be compatible with the DMN's
poorly understood involvement across
autobiographical memory recall, problem solving,
abstract reasoning, social cognition,
as well as delay discounting and self prospection into the future.
%
Improvement of the internal world representation
by injecting stochasticity into the recall of past
actions and the estimation of action outcomes may explain why
highly accurate memories have been disfavored in human evolution
and why human creativity may be adaptive.

A major hurdle in understanding DMN activity from cognitive neuroscienceå studies has been its similar engagement across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., daytime mind-wandering), and thinking about predicted scenarios (e.g., delay discounting).
The MDP account of DMN function offers a natural integration of a-priori diverging neural processes into a common framework.
%
It is an important advantage of the proposed artificial intelligence perspective
on DMN biology that it is practically computable and
readily motivates neuroscientific hypotheses that can be put to the test in future research.
%
Neuroscience experiments on the DMN should be designed that operationalize
the set of action, value, and state variables governing
the behavior of intelligent RL agents.
%
At the least, we propose an alternative vocabulary to
describe, contextualize, and interpret experimental findings in neuroscience studies
on higher-level cognition.
%
Ultimately,
neural processes in the DMN may instantiate a holistic integration
ranging from real experience over purposeful dreams to predicted futures
to continuously refine the organism's fate.
